Fast and scalable Gaussian process modeling with applications to astronomical time series \vspace{-3\baselineskip}

### 2017-02-15 paper/ms.tex
adding LAPACK_VARIANT option to build
adding sparse solver back in

To scale GP models to larger datasets, \citet{Rybicki:1995} presented a method
of computing \eq{gp-likelihood} in $\mathcal{O}(N)$ when the covariance
function is
\begin{eqnarray}\eqlabel{kernel-simple}
k_\bvec{\alpha}(\tau_{nm}) = \sigma_n^2\,\delta_{nm} + a\,\exp(-c\,\tau_{nm})
\end{eqnarray}
where $\{{\sigma_n}^2\}_{n=1}^N$ are the measurement uncertainties,
$\delta_{nm}$ is the Kronecker delta, and $\bvec{\alpha} = (a,\,c)$.
The intuition behind this method is that, for the choice of $k_\bvec{\alpha}$,
the inverse of $K_\bvec{\alpha}$ is tridiagonal and can it can be computed
with a small number of operations for each data point.
Subsequently, \citet{Ambikasaran:2015} generalized this method to arbitrary
mixtures of exponentials
\begin{eqnarray}
k_\bvec{\alpha}(\tau_{nm}) = \sigma_n^2\,\delta_{nm} +
    \sum_{j=1}^J a_j\,\exp(-c_j\,\tau_{nm})\quad.
\end{eqnarray}
In this case, the inverse will be dense but \eq{gp-likelihood} can still be
evaluated in $\mathcal{O}(J^2\,N)$ operations where $J$ is the number of
components in the mixture and $N$ is still the number of data points.

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.5\textwidth]{figures/carma_comp.pdf}
\caption{Benchmark\figurelabel{carma-comp}}
\end{center}
\end{figure}

>>>>

To scale GP models to larger datasets, \citet{Rybicki:1995} presented a method
of computing \eq{gp-likelihood} in $\mathcal{O}(N)$ when the covariance
function is
\begin{eqnarray}\eqlabel{kernel-simple}
k_\bvec{\alpha}(\tau_{nm}) = \sigma_n^2\,\delta_{nm} + a\,\exp(-c\,\tau_{nm})
\end{eqnarray}
where $\{{\sigma_n}^2\}_{n=1}^N$ are the measurement uncertainties,
$\delta_{nm}$ is the Kronecker delta, and $\bvec{\alpha} = (a,\,c)$.
The intuition behind this method is that, for the choice of $k_\bvec{\alpha}$,
the inverse of $K_\bvec{\alpha}$ is tridiagonal and can it can be computed
with a small number of operations for each data point.
Subsequently, \citet{Ambikasaran:2015} generalized this method to arbitrary
mixtures of exponentials
\begin{eqnarray}
k_\bvec{\alpha}(\tau_{nm}) = \sigma_n^2\,\delta_{nm} +
    \sum_{j=1}^J a_j\,\exp(-c_j\,\tau_{nm})\quad.
\end{eqnarray}
In this case, the inverse will be dense but \eq{gp-likelihood} can still be
evaluated in $\mathcal{O}(J^2\,N)$ operations \citep{Anderson:1999} where $J$
is the number of components in the mixture and $N$ is still the number of data
points.

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=\textwidth]{../examples/benchmark/benchmark_darwin_default.pdf}
\caption{Benchmark\figurelabel{benchmark-all}}
\end{center}
\end{figure}
